---
title: "Karina.Jonina.Project"
author: "KarinaJonina"
date: "5/1/2022"
output:
  html_document:
    df_print: paged
    toc: true
    theme: united
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE, warning=FALSE, message=FALSE}

library(tidyverse)
library(ggplot2)
library(dplyr)
library(stringr)
library(knitr)
library(plotly)
library(forcats)
library(crosstalk)
library(here)
library(camcorder)


DIRECTORY <- "data_2024"
# specifying the path to the CSV file
file_path <- here(DIRECTORY, "airbnb_final_df_20214.csv")

# Read the CSV file
df <- read_csv(file_path)


# changing 'month' to factor
df$month_category <- as.factor(df$month_category)
# changing  the order of factors
df$month_category <- factor(df$month_category, levels = c("1 month", 
                                                            "2 months",
                                                            "3 months",
                                                            "4 months",
                                                            "5 months",
                                                            "6 months",
                                                            "7 months",
                                                            "8 months",
                                                            "9 months",
                                                            "10 months",
                                                            "11 months",
                                                            "12 months",
                                                            "More than 12 months"
                                                            ))
# Print the first few rows to check
head(df)
```

```{r}
df %>%
  group_by(month_category) %>%
  summarise(n = n())
```


```{R echo=FALSE, warning=FALSE, message=FALSE}
camcorder::gg_record(dir = here("outputs"),
  device = "png",
  width = 8, height = 7,
  units = "in", dpi = 320
  )
```


### Final Dataset
In the final 'df' dataset, there are `r nrow(df) ` rows and `r ncol(df)` columns. Columns names are  `r colnames(df)`. 

Because our dataset was has been merged with 'available_nights_df', as described above, our dataset has AirBnB listings only available from 23rd October 2021 to 21st October 2022. Any past AirBnB are not included in our dataset. 

```{r echo=FALSE}
summary(df)
```

```{r echo=FALSE}
head(df)
```
```{r echo=FALSE}
df %>%
  select(id, county,room_type, availability_365, price, airbnb_potential_annual_income, #rent_price_per_month, 
         rent_potential_annual_income, best_choice_airbnb_or_rent_per_year,annual_diff_rent_minus_Airbnb) 
     
     
     
```
# Data Visualisation
```{r echo=FALSE}
# assigning colors to specific variable 
# will be used in graphs to ensure that the colours remain the same.
# creates association with colour and variables

#RPZ or Non-RPZ
pressure_colors = c( "#316395", "#D62728")

#Entire Unit, private room, shared room
room_colors = c('#54A24B', '#72B7B2', '#B279A2')

# rent or AirBnB
rent_colors = c('#565656', '#F58518')

#over-90 or under-90
threshold_colors = c('#FF9DA6', '#72B7B2')


# availability
reverse_viridis = c( "#FDE725FF",
                     '#B8DE29FF',
                     '#73D055FF',
                     '#29AF7FFF',
                     '#1F968BFF',
                     '#2D708EFF',
                     '#39568CFF',
                     '#482677FF',
                     "#440154FF")
```


# RPZ
```{r echo=FALSE, warning=FALSE}
# creating a map of Ireland with AirBnB listings in RTZ differentiated 
g <- list(
  scope = 'europe',
  showland = TRUE,
  landcolor = toRGB("gray95"),
  subunitcolor = toRGB("gray85"),
  countrycolor = toRGB("gray85"),
  countrywidth = 0.5,
  subunitwidth = 0.5,

  resolution = 50,
  lonaxis = list(range = c(-12, -5)),
  lataxis = list(range = c(52, 55))
)



unique_id_df <- df %>%
  group_by(id, rent_pressure_zone, lea, council, room_type, latitude, longitude) %>%
  summarise(price = mean(price))

fig <- plot_geo(df, lat = ~latitude, lon = ~longitude, color = ~rent_pressure_zone)
fig <- fig %>% add_markers(marker = list(opacity = 0.2), showlegend=FALSE, colors = pressure_colors,
                          
                          hovertemplate = ~paste("LEA:", lea, "<br>", 
                                        "Council:", council, "<br>",
                                        "Price (in €):", price, "<br>",
                                        "Room Type", room_type, "<br>",
                                        "RTZ", rent_pressure_zone, "<br>"))
fig <- fig %>% layout(
    title = 'Which AirBnB listings are in the RTZs?', geo = g)

fig
```


### Summary of RTZs
```{r message=FALSE, echo=FALSE}
# creating a dataframe with rent_pressure_zone and room_type, arranging by count
df %>% 
  select(rent_pressure_zone, room_type, price, minimum_nights) %>% 
  group_by(room_type, rent_pressure_zone) %>% 
  summarise(count = n(),
            min_price = min(price), 
            median_price = median(price), 
            mean_price = mean(price),
            max_price = max(price),
            min_nights = min(minimum_nights), 
            median_min_nights = median(minimum_nights),
            mean_min_nights = mean(minimum_nights),
            max_min_nights = max(minimum_nights)) 
```

```{r echo=FALSE, warning=FALSE}
# eamining the number of AirBnB listings available in each month
df %>%
  select(id, rent_pressure_zone) %>%
  group_by(rent_pressure_zone) %>%
  summarise(count_distinct = n_distinct(id))
```

```{r message=FALSE, echo=FALSE}
# creating a dataframe with rent_pressure_zone and room_type, arranging by count


df %>%
  group_by(rent_pressure_zone) %>%
  summarise(n = n_distinct(id)) %>%
  plot_ly(labels = ~rent_pressure_zone, values = ~n, type = 'pie',
          textinfo = 'label+percent',
          hoverinfo = 'label+value',
          hole = 0.5,
          marker = list(colors = pressure_colors)) %>%
  layout(title = list(text = '<b><span style="color:#D62728;">Over 11,300</span> of Airbnb listings are in <span style="color:#D62728;">Rent Pressure Zones</span>.</b>',
                      x = 0.5),
         showlegend = TRUE)

```

### Breakdown of AirBnB listings by Accomodation Type and RPZ
```{r echo=FALSE}
df %>%
  group_by(rent_pressure_zone, room_type) %>%
  summarise(n = n_distinct(id)) %>%
  mutate(room_type = fct_reorder(room_type, n, .desc = TRUE)) %>%
  plot_ly(x = ~room_type, y = ~n, color = ~rent_pressure_zone,
          
          hovertemplate = ~paste("RTZ:", rent_pressure_zone, "<br>",
                        "Accomodation Type:", room_type, "<br>", 
                        "Number of AirBnB listings:", n)) %>%
  add_bars(showlegend=FALSE, colors = pressure_colors) %>%
  layout(xaxis = list(title = FALSE),
          yaxis = list(title = "Number of AirBnB listings"),
         title = list(text = '<b>There are as Entire Units on Airbnb\n as Private Rooms in <span style="color:#D62728;">Rent Pressure Zones</span>.<b>',
                      x = 0.5))
   
```



```{r echo=FALSE}
df %>%
  group_by( rent_pressure_zone,best_choice_airbnb_or_rent_per_year) %>%
  summarise(n_id = n_distinct(id)) %>%
  plot_ly(x = ~rent_pressure_zone, y = ~n, color = ~best_choice_airbnb_or_rent_per_year) %>%
  add_bars(showlegend=FALSE, colors = rent_colors) %>%
  layout(xaxis = list(title = FALSE),
          yaxis = list(title = "Number of AirBnB listings"),
         title = list(text = '',
                      x = 0.5))
   
```

**Comment:**
Table above is summarised with this graph. There are more Entire Units and private rooms available outside of RTZs than there are in the RTZs. However, this could be due to fact that there are more houses in general outside the RTZs than inside the RTZs, which would show more AirBnB listings outside RTZs than inside RTZs.

### Breakdown of Price by: Room Type and non-/RTZ
```{r warning=FALSE, echo=FALSE}
df %>%
  plot_ly(x = ~room_type, y = ~price, color = ~rent_pressure_zone) %>%
  add_boxplot(marker = list(opacity = 0.2), showlegend = FALSE, colors = pressure_colors) %>% 
  layout(
    xaxis = list(title = FALSE),
    yaxis = list(title = "Price in Euro",
         zerolinecolor = '#ffff',
         zerolinewidth = 2,
         gridcolor = 'ffff',
         range = list(0, 450)), 
         title = list(text = '<b><span style="color:#D62728;">Entire Units in Rent Pressure Zoness</span> are\n on average <span style="color:#D62728;">more expensive</span> than units outside of RPZs\n or other accommodation types</b>',
                      x = 0.5),
    boxmode = "group") 
   
```

# Rent Pressure Zones

```{r echo=FALSE, warning=FALSE}

rpz_df <- df %>%
  filter(rent_pressure_zone == 'RPZ')
```


### Potential Annual Income
```{r echo=FALSE, warning=FALSE}
# geo styling
g <- list(
  scope = 'europe',
  showland = TRUE,
  landcolor = toRGB("gray95"),
  subunitcolor = toRGB("gray85"),
  countrycolor = toRGB("gray85"),
  countrywidth = 0.5,
  subunitwidth = 0.5,

  resolution = 50,
  lonaxis = list(range = c(-12, -5)),
  lataxis = list(range = c(52, 55))
)

fig <- plot_geo(rpz_df, lat = ~latitude, lon = ~longitude, color = ~best_choice_airbnb_or_rent_per_year)
fig <- fig %>% add_markers(marker = list(opacity = 0.4),colors = rent_colors,
                hovertemplate = ~paste("County:", county, "<br>",
                        "To Rent or To Buy:",best_choice_airbnb_or_rent_per_year,"<br>",
                        
                        "Potential Difference", annual_diff_rent_minus_Airbnb, "<br>",
                        
                        "Airbnb Potential Annual Income:", airbnb_potential_annual_income, "<br>",
                        "Rent Potential Annual Income:", rent_potential_annual_income
                        ))
fig <- fig %>% layout(
    title = list(
        text = '<b>Based on Potential Annual Income,\n<span style="color:#F58518;">Long-Term Leasing</span> would potentially bring as much as Airbnb</b>',
        x = 0.5,
        y = 0.95,
        yanchor = "top"
    ),
    legend = list(
        orientation = "h",
        y = -0.1,
        x = 0.5,
        xanchor = "center",
        yanchor = "bottom"
    ),
    margin = list(
        l = 50,  # left margin
        r = 50,  # right margin
        t = 100, # top margin
        b = 50   # bottom margin
    ),
    geo = g
)

fig
```



```{r echo=FALSE, warning=FALSE}

rpz_df %>%
  select(id, best_choice_airbnb_or_rent_per_year) %>%
  group_by(best_choice_airbnb_or_rent_per_year) %>%
    summarise(n = n_distinct(id))
```

```{r message=FALSE, echo=FALSE}
rpz_df %>%
  group_by(best_choice_airbnb_or_rent_per_year) %>%
    summarise(n = n_distinct(id)) %>%
  plot_ly(labels = ~best_choice_airbnb_or_rent_per_year, values = ~n, type = 'pie',
          textinfo = 'label+percent',
          hoverinfo = 'label+value',
          hole = 0.5,
          marker = list(colors = rent_colors)) %>%
  layout(title = list(text = '<b><span style="color:#F58518;">Over 4,400 </span> landlords would earn <span style="color:#F58518;">more income</span>\n by <span style="color:#F58518;">Long-Term Leasing</span> than Airbnb in RPZs.</b>',
                      x = 0.5),

         
         
        legend = list(
        orientation = "h",
        y = -0.1,
        x = 0.5,
        xanchor = "center",
        yanchor = "bottom"
    ),
    margin = list(
        l = 50,  # left margin
        r = 50,  # right margin
        t = 100, # top margin
        b = 50   # bottom margin
    ))

```

## Splitting data by room_type
A new dataset was created where AirBnB listings were selected that offered Entire Unit.  This new dataset was used for this section of the report.  
```{r echo=FALSE}
# create new dataset where filtered by 'Entire Unit'
entire_rpz <-  rpz_df %>% 
  filter(room_type =='Entire rental unit')



entire_rpz_count <- entire_rpz %>%
  group_by(best_choice_airbnb_or_rent_per_year) %>%
    summarise(n = n_distinct(id)) 


#######################
# annual_diff_rent_minus_Airbnb
#######################

# If POSITIVE, then AIRBNB brings more income
# If NEGATIVE, then RENT brings more income

entire_rpz_more_income_rent <-  entire_rpz %>%
  filter(annual_diff_rent_minus_Airbnb <0) %>%
  mutate(annual_diff_rent_minus_Airbnb = abs(annual_diff_rent_minus_Airbnb))


private_room_rpz_more_income_rent <-  private %>%
  filter(annual_diff_rent_minus_Airbnb <0)  %>%
  mutate(annual_diff_rent_minus_Airbnb = abs(annual_diff_rent_minus_Airbnb))

#######################
#
#######################
# create new dataset where filtered by 'private 
private <-  rpz_df %>% 
  filter(room_type =='Private room')


private_rpz_count <- private %>%
  group_by(best_choice_airbnb_or_rent_per_year) %>%
  summarise(n = n_distinct(id))






```

```{r echo=FALSE}


fig <- plot_ly()
fig <- fig %>% add_pie(data = entire_rpz_count, labels = ~best_choice_airbnb_or_rent_per_year, values = ~n,
                      name = "Entire Unit", 
                      textinfo = 'label+percent+name',
                      hoverinfo = 'label+value+name',
                      marker = list(colors = rent_colors),
                      hole = 0.45,
                      domain = list(row = 0, column = 0))

fig <- fig %>% add_pie( data = private_rpz_count, labels = ~best_choice_airbnb_or_rent_per_year, values = ~n,
                      name = "Private Room", 
                      textinfo = 'label+percent+name',
                      hoverinfo = 'label+value+name',
                      marker = list(colors = rent_colors),
                      hole = 0.45, domain = list(row = 0, column = 1))

fig <- fig %>% layout(title = '<b><span style="color:#F58518;">Over One-Third Entire Units and Private Rooms</span>\nwould potentially bring more income by <span style="color:#F58518;">Long-Term Leasing</span> than Airbnb in RPZs.</b>',
              showlegend = F,
              grid=list(rows=1, columns=2),
              xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
              yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
              
              annotations = list(
                list(
                text = '<b>Entire Unit</b>',
                x = 0.15,
                y = 0,
                xref = 'paper',
                yref = 'paper',
                showarrow = FALSE
                ),
                            list(
              text = '<b>Private Room</b>',
              x = 0.85,
              y = 0,
              xref = 'paper',
              yref = 'paper',
              showarrow = FALSE
              )
              )
              )

fig
```

## Entire Rental

```{r echo=FALSE}
entire_rpz_more_income_rent %>%
  group_by(county) %>%
  summarise(n_id = n_distinct(id)) %>%
  mutate(county = fct_reorder(county, n_id, .desc = T)) 

```


```{r echo=FALSE}

library(forcats)


# Create the plot
fig <- entire_rpz_more_income_rent %>%
  group_by(county, best_choice_airbnb_or_rent_per_year) %>%
  summarise(n_id = n_distinct(id)) %>%
  ungroup() %>%  # Ensure the factor reordering is applied correctly
  mutate(county = fct_reorder(county, n_id, .desc = FALSE)) %>%
  plot_ly(y = ~county, x = ~n_id, color = ~best_choice_airbnb_or_rent_per_year,
          hovertemplate = ~paste("County:", county, "<br>",
                                 "Accomodation Type:", best_choice_airbnb_or_rent_per_year, "<br>", 
                                 "Number of AirBnB listings:", n_id)) %>%
  add_bars(showlegend = FALSE, marker = list(color = '#F58518')) %>%
  layout(
    yaxis = list(title = FALSE, categoryorder = "total ascending"),  # Order y-axis by n_id
    xaxis = list(title = "Number of AirBnB listings that would bring more income by Long-Term Leasing than Airbnb"),
    title = list(
      text = '<b>Based on Potential Annual Income,\nnearly 1,400 Entire Units would bring more income\nby <span style="color:#F58518;">Long-Term Leasing</span> in Dublin</b>',
      x = 0.5,
      y = 0.95,
      xanchor = 'center',
      yanchor = 'top'
    ),
    margin = list(t = 100, b = 50)  # Adjust margins to fit the title and graph
  )

fig

```


```{r warning=FALSE, echo=FALSE}
entire_rpz_more_income_rent_month_category_county_table <- entire_rpz_more_income_rent %>%
  group_by(month_category, county, id) %>%
  
  
  ungroup() %>%
   group_by(month_category, county)%>%
   
  summarise(
    n = n(),
    min_profit = min(annual_diff_rent_minus_Airbnb),
    mean_profit = mean(annual_diff_rent_minus_Airbnb), 
    median_profit = median(annual_diff_rent_minus_Airbnb),  
    max_profit = max(annual_diff_rent_minus_Airbnb),
            ) 
print(entire_rpz_more_income_rent_month_category_county_table)
```

```{r warning=FALSE, echo=FALSE}

crosstable_function <- function(table) { # create a function with the name my_function
  entire_month_category_county_crosstable <- table %>%
    select(month_category, county, mean_profit) %>%
    pivot_wider(names_from = month_category, values_from = mean_profit)
  
  
  
  # Order by the '1 month' column in descending order
  entire_month_category_county_crosstable <- entire_month_category_county_crosstable %>%
    arrange(desc(`1 month`), desc = FALSE) %>%
    mutate(across(where(is.numeric), ~ round(., -1)))
  
  entire_month_category_county_crosstable[is.na(entire_month_category_county_crosstable)] <- 0
  
  # Create a heatmap
  heatmap_data <- as.matrix(select(entire_month_category_county_crosstable, -county))
  rownames(heatmap_data) <- entire_month_category_county_crosstable$county
  
  
  heatmap <- plot_ly(
    x = colnames(heatmap_data),
    y = rownames(heatmap_data),
    z = heatmap_data,
    type = "heatmap",
    colors = colorRamp(c("white", "#F58518"))
  ) %>%
    layout(
      title = "Average Potential Annual Earning Difference of Entire Units in RPZs by Month Category and County",
      xaxis = list(title = ""),
      yaxis = list(title = "")
    )
  
  
  # Add annotations
  for (i in seq_len(nrow(heatmap_data))) {
    for (j in seq_len(ncol(heatmap_data))) {
      heatmap <- heatmap %>%
        add_annotations(
          x = colnames(heatmap_data)[j],
          y = rownames(heatmap_data)[i],
          text = round(heatmap_data[i, j], 2),
          showarrow = FALSE,
          font = list(color = 'black')
        )
    }
  }
  
  # Show the plot
  return(heatmap)

}



crosstable_function(entire_rpz_more_income_rent_month_category_county_table)

```





```{r}
entire_rpz_more_income_rent %>%
  plot_ly(x = ~month_category, y = ~annual_diff_rent_minus_Airbnb) %>%
  add_boxplot(marker = list(opacity = 0.2), showlegend = FALSE, colors = '#565656') %>% 
  layout(
    xaxis = list(title = FALSE),
    yaxis = list(title = "Difference between Airbnb and Long-Term Leasing"#,
         # zerolinecolor = '#ffff',
         # zerolinewidth = 2,
         # gridcolor = 'ffff',
         # range = list(0, 450)
         ), 
         title = list(text = '<b>Landlords <span style="color:#565656;">Airbnb</span> short number of months make less money than those who Long-Term Lease</b> Earnings',
                      x = 0.5),
    boxmode = "group") 
   
```


## private Room


```{r echo=FALSE}
private_room_rpz_more_income_rent %>%
  group_by(county, best_choice_airbnb_or_rent_per_year) %>%
  summarise(n_id = n_distinct(id)) %>%
  ungroup() %>%  # Ensure the factor reordering is applied correctly
  mutate(county = fct_reorder(county, n_id, .desc = FALSE)) %>%
  plot_ly(y = ~county, x = ~n_id, color = ~best_choice_airbnb_or_rent_per_year,
          hovertemplate = ~paste("County:", county, "<br>",
                                 "Accomodation Type:", best_choice_airbnb_or_rent_per_year, "<br>", 
                                 "Number of AirBnB listings:", n_id)) %>%
  add_bars(showlegend = FALSE, marker = list(color = '#F58518')) %>%
  layout(
    yaxis = list(title = FALSE, categoryorder = "total ascending"),  # Order y-axis by n_id
    xaxis = list(title = "Number of AirBnB listings that would bring more income by Long-Term Leasing than Airbnb"),
    title = list(
      text = '<b>Based on Potential Annual Income,\nnearly 700 Private Rooms would bring more income\nby <span style="color:#F58518;">Long-Term Leasing</span> in Dublin</b>',
      x = 0.5,
      y = 0.95,
      xanchor = 'center',
      yanchor = 'top'
    ),
    margin = list(t = 100, b = 50)  # Adjust margins to fit the title and graph
  )

   
```

```{r warning=FALSE, echo=FALSE}
private_room_month_category_county_table <- private_room_rpz_more_income_rent %>%
  group_by(month_category, county) %>%
  summarise(
    min_profit = min(annual_diff_rent_minus_Airbnb),
    mean_profit = mean(annual_diff_rent_minus_Airbnb), 
    median_profit = median(annual_diff_rent_minus_Airbnb),  
    max_profit = max(annual_diff_rent_minus_Airbnb),
            ) 


private_room_month_category_county_crosstable <- private_room_month_category_county_table %>%
  select(month_category, county, mean_profit) %>%
  pivot_wider(names_from = month_category, values_from = mean_profit)



# Order by the '1 month' column in descending order
private_room_month_category_county_crosstable <- private_room_month_category_county_crosstable %>%
  arrange(desc(`1 month`), desc = FALSE) %>%
  mutate(across(where(is.numeric), ~ round(., -1)))

private_room_month_category_county_crosstable[is.na(private_room_month_category_county_crosstable)] <- 0

# Create a heatmap
heatmap_data <- as.matrix(select(private_room_month_category_county_crosstable, -county))
rownames(heatmap_data) <- private_room_month_category_county_crosstable$county


heatmap <- plot_ly(
  x = colnames(heatmap_data),
  y = rownames(heatmap_data),
  z = heatmap_data,
  type = "heatmap",
  colors = colorRamp(c("white", "#F58518"))
) %>%
  layout(
    title = "Heatmap of Mean Profit by Month Category and County",
    xaxis = list(title = "Month Category"),
    yaxis = list(title = "County")
  )


# Add annotations
for (i in seq_len(nrow(heatmap_data))) {
  for (j in seq_len(ncol(heatmap_data))) {
    heatmap <- heatmap %>%
      add_annotations(
        x = colnames(heatmap_data)[j],
        y = rownames(heatmap_data)[i],
        text = round(heatmap_data[i, j], 2),
        showarrow = FALSE,
        font = list(color = 'black')
      )
  }
}

# Show the plot
heatmap

```

```{r warning=FALSE, echo=FALSE}
private_room_rpz_more_income_rent %>%
  plot_ly(x = ~month_category, y = ~annual_diff_rent_minus_Airbnb) %>%
  add_boxplot(marker = list(opacity = 0.2), showlegend = FALSE, colors = '#565656') %>% 
  layout(
    xaxis = list(title = FALSE),
    yaxis = list(title = "Difference between Airbnb and Long-Term Leasing"#,
         # zerolinecolor = '#ffff',
         # zerolinewidth = 2,
         # gridcolor = 'ffff',
         # range = list(0, 450)
         ), 
         title = list(text = '<b>Landlords <span style="color:#565656;">Airbnb</span> short number of months make less money than those who Long-Term Lease</b> Earnings',
                      x = 0.5),
    boxmode = "group") 
   
```


## AirBnB OverView
```{r echo=FALSE, warning=FALSE}
# creating a map of Ireland with AirBnB listings
g <- list(
  scope = 'europe',
  showland = TRUE,
  landcolor = toRGB("gray95"),
  subunitcolor = toRGB("gray85"),
  countrycolor = toRGB("gray85"),
  countrywidth = 0.5,
  subunitwidth = 0.5,

  resolution = 50,
  lonaxis = list(range = c(-12, -5)),
  lataxis = list(range = c(52, 55))
)

fig <- plot_geo(df, lat = ~latitude, lon = ~longitude, color = ~county)
fig <- fig %>% add_markers(marker = list(opacity = 0.4), showlegend = FALSE,
                
                hovertemplate = ~paste("County:", county, "<br>", 
                        "Price (in €):", price, "<br>",
                        "Room Type", room_type, "<br>",
                        "Mininum stay (in nights)", minimum_nights, "<br>",
                        "Availability (in nights)", available_nights))
fig <- fig %>% layout(
    title = 'Where are the AirBnB listings in Ireland?', geo = g)

fig
```
**Comment:**
It is clear from this graph that there are some parts of Ireland that have more AirBnB listings than others. It is easy to see that there are more properties available on the coast than inland. 


### Room Types are available on AirBnB
``` {r}
df %>%
  group_by(room_type) %>%
    summarise(n = n_distinct(id)) %>%
  plot_ly(labels = ~room_type, values = ~n, type = 'pie',
          textinfo = 'label+percent',
          hoverinfo = 'label+value',
          marker = list(colors =room_colors),
          hole = 0.45) %>%
  layout(title = list(text = '<b>The Most Common Acccomodation Type in Ireland is <span style="color:#54A24B;">Entire Units</span></b>',
                      x = 0.5),
         showlegend = FALSE)

```

### Where are the AirBnB listings in Ireland?
```{r warning=FALSE, echo=FALSE}
df %>%
  group_by(county) %>%
    summarise(n = n_distinct(id)) %>%
  mutate(county = fct_reorder(county, n, .desc = FALSE)) %>%
  plot_ly(y = ~county, x = ~n,
          
          hovertemplate = ~paste("County:", county, "<br>", "Number of AirBnB listings:", n)) %>%
  add_bars(showlegend = FALSE) %>%
  layout(yaxis = list(title = FALSE),
          xaxis = list(title = "Number of AirBnB listings"),
         title= 'Which counties have the most AirBnB listings in Ireland?') 
```

### When are most AirBnB listings Available?



```{r echo=FALSE, warning=FALSE}
# specifying the path to the CSV file
file_path <- here(DIRECTORY, "airbnb_large_df_2024.csv")

# Read the CSV file
large_df <- read_csv(file_path)


# changing 'month' to factor
large_df$month <- as.factor(large_df$month)
# changing  the order of factors
large_df$month <- factor(large_df$month, levels = c("August 2024",
                                                          "September 2024",
                                                           "October 2024",
                                                          "November 2024",
                                                          "December 2024",
                                                          "January 2025",
                                                          "February 2025",
                                                          "March 2025",
                                                          "April 2025",
                                                          "May 2025",
                                                          "June 2025",
                                                          "July 2024",
                                                          "June 2024"   
                                                            ))

```


```{r echo=FALSE, warning=FALSE}

sum_df <- large_df %>% 
  select(id, month) %>% 
  group_by(month) %>% 
  summarize(count_distinct = n_distinct(id))


sum_df %>%
  plot_ly(x = ~month, y = ~count_distinct, #color = ~rent_pressure_zone,
          hovertemplate = ~paste("Month:", month, "<br>", 
                                 "Number of AirBnB listings:", count_distinct)) %>%
  add_lines(showlegend = FALSE) %>%
  layout(xaxis = list(title = FALSE),
         yaxis = list(title = "Number of AirBnB listings", range = c(0, max(sum_df$count_distinct))),
         title = "When are the most AirBnB listings Available?")

```

